Architectures and Open pNets

### 2018-04-06 2018-Arch-pNets/arch-pNets.tex
Updated the proofs to include maximal progress Had to introduce the notion of upwards compatibility

\newcommand{\defn}[1]{Def.~\ref{defn:#1}}
\newcommand{\defntwo}[2]{Defs~\ref{defn:#1} and \ref{defn:#2}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\figs}[2]{Fig.~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\tab}[1]{Tab.~\ref{tab:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\res}[1]{(\ref{res:#1})}
\newcommand{\ex}[1]{Ex.~\ref{ex:#1}}
\newcommand{\exs}[2]{Ex.~\ref{ex:#1} and~\ref{ex:#2}}
\newcommand{\secn}[1]{Sect.~\ref{secn:#1}}
\newcommand{\rem}[1]{Rem.~\ref{rem:#1}}
\newcommand{\lem}[1]{Lem.~\ref{lem:#1}}
\newcommand{\cor}[1]{Cor.~\ref{cor:#1}}
\newcommand{\thm}[1]{Th.~\ref{thm:#1}}
\newcommand{\axs}[1]{Ax.~\ref{ax:#1}}
% \newcommand{\ax}[2]{\ref{ax:#1}.\ref{ax:#1:#2}}
\newcommand{\ax}[1]{Ax.~\ref{ax:#1}}
\newcommand{\prop}[1]{Prop.~\ref{prop:#1}}
\newcommand{\alg}[1]{Alg.~\ref{alg:#1}}
\newcommand{\asmp}[1]{Ass.~\ref{asmp:#1}}

@@

  In fact, I am not sure whether I have correctly understood this
  suggestion: do you mean, \eg using $g$ to denote the set of
  Boolean expressions or individual Boolean expressions? Using
  lower case letters to denote sets might be problematic, since
  we (at least I) use them a lot to denote instances, as in
  \[q \goesto{a, g, e} q'\,.\]
  In fact, I even use capital letters: in the definition of
  interaction model composition, I use
  \[\cdot \goesto{a, G, E} \cdot\]
  in the conclusion of the rule.  However, for this latter I can
  easily change the notation.

@@

  \begin{tabular}{
      @{}p{0.15\columnwidth}@{}
      p{0.35\columnwidth}@{}
      p{0.5\columnwidth}@{}
    }
    {\bf Output} & {\bf Command} & {\bf Definition}
    \\
    $\data$ &
    {\ttfamily\textbackslash data} &
    {\ttfamily\textbackslash sD}
    \\
    \signature &
    {\ttfamily\textbackslash signature} &
    {\ttfamily\textbackslash Sigma}
    \\
    \variables &
    {\ttfamily\textbackslash variables} &
    {\ttfamily\textbackslash cV}
    \\
    \Talg &
    {\ttfamily\textbackslash Talg} &
    {\ttfamily\textbackslash cT\_\{\textbackslash signature,\textbackslash variables\}}
    \\
    \actions{\#1} &
    {\ttfamily\textbackslash actions\{\#1\}} &
    {\ttfamily\textbackslash cA\_\{\#1\}}
    \\
    \exprs{\#1} &       
    {\ttfamily\textbackslash exprs\{\#1\}} &
    {\ttfamily\textbackslash sE\_\{\#1\}}
    \\
    \boolexprs{\#1} &       
    {\ttfamily\textbackslash boolexprs\{\#1\}} &
    {\ttfamily\textbackslash sB\_\{\#1\}}
    \\
    \guards{\#1} &
    {\ttfamily\textbackslash guards\{\#1\}} &
    {\ttfamily\textbackslash sB\^{}\{<\}\_\{\#1\}}
    \\
    \assigns{\#1} &
    {\ttfamily\textbackslash assigns\{\#1\}} &
    {\ttfamily\textbackslash sA\_\{\#1\}}
    \\
    \valuations{\#1} &
    {\ttfamily\textbackslash valuations\{\#1\}} &
    {\ttfamily\textbackslash data\^{}\{\#1\}}
    \\[6pt]
    \val{\#2}{\#3} &
    {\ttfamily\textbackslash val\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
    \\[6pt]
    \val[\tilde]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash tilde]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
  \end{tabular}
\noteSBin{commands above (tabular generates an error inside a
  note)}

@@

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, V, \val{0}{}, P,
  \export, \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    \noteSB{Added this}\hl{$\export : P \rightarrow 2^V$ 
    the \emph{export function},}
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \exprs{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports
    $\emptyset \neq a \subseteq P$, a monotonic Boolean \emph{guard} $g \in
    \guards{V}$ and a monotonic \emph{update expression} \todoSB{$e \rightarrow u$}\hl{$e \in
    \exprs{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, component variables are split in two sub-sets:
  local and exported variables.  Only exported variables belong
  to the component interface and can be used for the interactions
  with other components (see \defn{im}).  However, in the context
  of this paper we can simplify by omitting this separation.
%
  }
%
  Notations $q \goesto{a} q'$ and $q \goesto{a}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

@@

  We denote $\filter{\alpha} \bydef{=} \setdefb{\beta \in (2^P \times
    \guards{V} \times \exprs{V})}{\alpha \prec_\pi \beta}$ the set of
  interactions that have higher priority than $\alpha$.
\end{definition}

\begin{definition}[Application of priority]
  \label{defn:priority:sem}
  The application of a priority model $\pi$ over $(V, P, \export)$
  to a component $B = (Q, q^0, V, \val{0}{}, P, \export, \goesto{})$
  is the component $\pi(B) = (Q, q^0, V, \val{0}{}, P, \export,
  \goesto[\pi]{})$, where $\goesto[\pi]{}$ is the minimal transition
  relation satisfying the rule
%
  \begin{equation}
    \label{eq:priority:sem}
%
    \infer{    
      q \goesto{a, g, e} q'
      \and
      \textstyle
      g' = g \land \bigwedge_{(b, h, f) \in \filter[q]{\alpha}} \lnot h
    }{
      q \goesto{a, g', e} q'
    }\,, 
  \end{equation}
%
  where $\filter[q]{\alpha} \bydef{=} \filter{\alpha} \cap
  \setdefb{\beta \in (2^P \times \guards{V} \times \exprs{V})}{q
    \goesto{\beta}}$.
  \todoSBin{Add the intuitive explanation}
\end{definition}

@@

\begin{note}
  Below, we will assume that the maximal progress priority is
  applied after each composition with an interaction model.
\end{note}

@@

\begin{definition}[Application of an architecture]
  \label{defn:arch:application}
  Let $A = (\cC, V_A, P_A, \export[A], \Gamma)$ be an architecture and let $\cB$
  be a set of components, such that
%
  \begin{align}
    \bigcup_{B \in \cB} V_B \cap \bigcup_{C \in \cC} V_C = \emptyset\,,
    &&
    V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B\,,
    \\
    \bigcup_{B \in \cB} P_B \cap \bigcup_{C \in \cC} P_C = \emptyset\,,
    &&
    P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B\,,
  \end{align}
%
  \hl{and $\export[A](p) = V_A \cap \export[B](p)$, for any $B \in \cB$
  and $p \in P_A \cap P_B$.}
%
  The \emph{application of the architecture $A$ to the set of
  components $\cB$} is the component $ A(\cB) \bydef{=}
  \mu\bigl((\IMextend{\Gamma}{P})(\cC \cup \cB)\bigr)$, where
%
  \begin{equation}
    \label{eq:im:extension}
    \IMextend{\Gamma}{P} \bydef{=}
    \setdefb{
      (a \cup a', g, e)
    }{
      (a, g, e) \in \Gamma, a' \subseteq P \setminus P_A
    }
  \end{equation}
%
  denotes the \emph{extension} of the interaction model $\Gamma$
  to the set of ports $P$
  and $\mu$ is the maximal progress priority model.
\end{definition}

@@

\begin{definition}[Composition of architectures]
  \label{defn:arch:composition}
  Let $A_i = (\cC_i, V_{A_i}, P_{A_i}, \export[A_i], \Gamma_i)$, for $i = 1,2$,
  be two architectures.  The \emph{composition} of $A_1$ and
  $A_2$ is the architecture $A_1 \arcomp A_2 = (\cC_1 \cup \cC_2,
  V_{A_1} \cup V_{A_2}, P_{A_1} \cup P_{A_2}, \export[A_1] \cup \export[A_2], \Gamma)$, where
%
  \begin{equation}
    \label{eq:arch:composition}
    \Gamma = \setdefb{
      (a, g_1 \land g_2, e_1 \expmix e_2) 
    }{
      (a \cap P_{A_i}, g_i, e_i) \in \Gamma_i,
      \text{ for } i = 1,2
    }
    \,.
  \end{equation}
\end{definition}

\begin{proposition}[Properties of $\arcomp$]
  \label{prop:arcomp:nice}
  Architecture composition $\arcomp$ is commutative and
  associative; it is idempotent if all coordinating components
  are deterministic; $A_{id} = \bigl(\emptyset, \emptyset,
  \emptyset, \{\emptyset\}\bigr)$ is its neutral element, \ie for
  any architecture $A$, we have $A \oplus A_{id} \arequiv A$.
  Furthermore, for any component $B$, we have $A_{id}(B) = B$.
\end{proposition}
%
\begin{proof}[Sketch of the proof]
  Commutativity and associativity follow from the corresponding
  properties of set union, Boolean conjunction and
  \todoSB{Check ``semi''}\hl{semilattice meet}.
%
  Suppose we have two architectures $A \arequiv A'$.  This does not
  necessarily mean that their sets of coordinating components
  coincide.  However, if all the involved coordinating components
  are deterministic, then, in any state of $(A \arcomp A')(\cB)$,
  both architectures will impose the same restrictions, enabling
  the same interactions between the coordinating and operand
  components.  Hence, we have $(A \arcomp A')(\cB) = A(\cB) =
  A'(\cB)$.  Since this holds for any set of components $\cB$, we
  conclude that $A \arcomp A' \arequiv A \arequiv A'$.
%
  The properties of $A_{id}$ follow immediately from the
  definitions of architecture application and composition.
\end{proof}

@@

\begin{theorem}[Preserving enforced properties]
  \label{thm:combining}
  Let $\cB$ be a set of components; let $A_i = (\cC_i, V_{A_i},
  P_{A_i}, \export[A_i], \Gamma_i)$, for $i = 1,2$, be two architectures
  enforcing on $\cB$ the properties $\Phi_1$ and $\Phi_2$
  respectively.  The composition $A_1 \arcomp A_2$ enforces on
  $\cB$ the property $\Phi_1 \land \Phi_2$.
\end{theorem}

@@

\begin{lemma}
  \label{lem:onlyone}
  Let $A = (\cC, V_A, P_A, \Gamma)$ be an architecture and denote
  by $\Gamma_\cC \bydef{=}
%
  \setdef{
    (a \cap P_\cC, \true, \noop)
  }{
    (a, g, e) \in \Gamma
  }$, with $P_\cC = \bigcup_{C \in \cC} P_C$,
%  
  the projection of $\Gamma$ onto the coordinating components of
  $A$.  Consider the architecture $A' = (\{C'\}, V_A, P_A,
  \Gamma)$, where $C' = \Gamma_\cC(\cC)$.  For any set of
  components $\cB$, satisfying the conditions of
  \defn{arch:application}, we have
  $\semopen{A(\cB)} = \semopen{A'(\cB)}$.
\end{lemma}
%
\begin{proof}
  First of all, notice that, by \defn{im},
  $V_{C'} = \bigcup_{C \in \cC} V_C$ and $P_{C'} = P_\cC$.
  The conditions of \defn{arch} are satisfied and $A'$ is indeed
  an architecture.  Furthermore, $\cB$ satisfies the conditions
  of \defn{arch:application} \wrt $A'$.  Hence, the component
  $A'(\cB)$ is well defined.

  Clearly the state spaces, initial states and interfaces of both
  components coincide.  Thus we only have to prove that so do the
  transition relations.  Let us assume that
  $\cC = C_i^{i \in I}$ and $\cB = B_j^{j \in J}$.
  We will use $q_i, q_i'$ to denote the states of
  $C_i$ and $q_j, q_j'$ to denote the states of $B_j$,
  and similarly for the valuations of variables.

  By \defn{comp:semantics}, a transition
%
  \begin{equation}
    \label{eq:lem1:trans:sem}
    (q_i, \val{}{i})^{i \in I} (q_j, \val{}{j})^{j \in J}
    \goesto{a, \val[\tilde]{}{}}
    (q_i', \val{\prime}{i})^{i \in I}
    (q_j', \val{\prime}{j})^{j \in J}
  \end{equation}
%  
  is a transition in $\semopen{A(\cB)}$ (\resp
  $\semopen{A'(\cB)}$) iff
%
  \begin{equation}
    \label{eq:lem1:trans}
    q_i^{i \in I} q_j^{j \in J}
    \goesto{a, G, E}
    (q_i')^{i \in I} (q_j')^{j \in J}
    \,,
  \end{equation}
%
  is a transition in $A(\cB)$ (\resp $A'(\cB)$) and
%
  \begin{mathpar}
    \val{i \in I}{i} \val{j \in J}{j}
    \models G
    \,,
    \and
    (\val{\prime}{i})^{i \in I} (\val{\prime}{j})^{j \in J}
    = \val[\tilde]{}{}[E]
    \,,
    \and
    \valdiff{\val{i \in I}{i} \val{j \in J}{j}}{
      \val[\tilde]{}{}} \subseteq \export(a)
    \,.   
  \end{mathpar}

>>>>

\newcommand{\defn}[1]{Def.~\ref{defn:#1}}
\newcommand{\defntwo}[2]{Defs~\ref{defn:#1} and \ref{defn:#2}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\figs}[2]{Fig.~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\tab}[1]{Tab.~\ref{tab:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\res}[1]{(\ref{res:#1})}
\newcommand{\ex}[1]{Ex.~\ref{ex:#1}}
\newcommand{\exs}[2]{Ex.~\ref{ex:#1} and~\ref{ex:#2}}
\newcommand{\secn}[1]{Sect.~\ref{secn:#1}}
\newcommand{\rem}[1]{Note~\ref{rem:#1}}
\newcommand{\lem}[1]{Lem.~\ref{lem:#1}}
\newcommand{\cor}[1]{Cor.~\ref{cor:#1}}
\newcommand{\thm}[1]{Th.~\ref{thm:#1}}
\newcommand{\axs}[1]{Ax.~\ref{ax:#1}}
% \newcommand{\ax}[2]{\ref{ax:#1}.\ref{ax:#1:#2}}
\newcommand{\ax}[1]{Ax.~\ref{ax:#1}}
\newcommand{\prop}[1]{Prop.~\ref{prop:#1}}
\newcommand{\alg}[1]{Alg.~\ref{alg:#1}}
\newcommand{\asmp}[1]{Ass.~\ref{asmp:#1}}

@@

  In fact, I am not sure whether I have correctly understood this
  suggestion: do you mean, \eg using $g$ to denote the set of
  Boolean expressions or individual Boolean expressions? Using
  lower case letters to denote sets might be problematic, since
  we (at least I) use them a lot to denote instances, as in
  \[q \goesto{a, g, e} q'\,.\]
  %% In fact, I even use capital letters: in the definition of
  %% interaction model composition, I use
  %% \[\cdot \goesto{a, G, E} \cdot\]
  %% in the conclusion of the rule.  However, for this latter I can
  %% easily change the notation.

@@

  \begin{tabular}{
      @{}p{0.15\columnwidth}@{}
      p{0.35\columnwidth}@{}
      p{0.5\columnwidth}@{}
    }
    {\bf Output} & {\bf Command} & {\bf Definition}
    \\
    $\data$ &
    {\ttfamily\textbackslash data} &
    {\ttfamily\textbackslash sD}
    \\
    \signature &
    {\ttfamily\textbackslash signature} &
    {\ttfamily\textbackslash Sigma}
    \\
    \variables &
    {\ttfamily\textbackslash variables} &
    {\ttfamily\textbackslash cV}
    \\
    \Talg &
    {\ttfamily\textbackslash Talg} &
    {\ttfamily\textbackslash cT\_\{\textbackslash signature,\textbackslash variables\}}
    \\
    \actions{\#1} &
    {\ttfamily\textbackslash actions\{\#1\}} &
    {\ttfamily\textbackslash cA\_\{\#1\}}
    \\
    \exprs{\#1} &       
    {\ttfamily\textbackslash exprs\{\#1\}} &
    {\ttfamily\textbackslash sE\_\{\#1\}}
    \\
    \boolexprs{\#1} &       
    {\ttfamily\textbackslash boolexprs\{\#1\}} &
    {\ttfamily\textbackslash sB\_\{\#1\}}
    \\
    \guards{\#1} &
    {\ttfamily\textbackslash guards\{\#1\}} &
    {\ttfamily\textbackslash sB\^{}\{<\}\_\{\#1\}}
    \\
    \assigns{\#1} &
    {\ttfamily\textbackslash assigns\{\#1\}} &
    {\ttfamily\textbackslash sA\_\{\#1\}}
    \\
    \valuations{\#1} &
    {\ttfamily\textbackslash valuations\{\#1\}} &
    {\ttfamily\textbackslash data\^{}\{\#1\}}
    \\[6pt]
    \val[]{\#2}{\#3} &
    {\ttfamily\textbackslash val[]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
    \\[6pt]
    \val[\tilde]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash tilde]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
  \end{tabular}
\noteSBin{commands above (tabular generates an error inside a
  note)}

@@

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, V, \val{0}{}, P,
  \export, \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    \noteSB{Added this}\hl{$\export : P \rightarrow 2^V$ 
    the \emph{export function},}
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \exprs{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports
    $\emptyset \neq a \subseteq P$, a monotonic Boolean \emph{guard} $g \in
    \guards{V}$ and a monotonic \emph{update expression} \todoSB{$e \rightarrow u$}\hl{$e \in
    \exprs{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, component variables are split in two sub-sets:
  local and exported variables.  Only exported variables belong
  to the component interface and can be used for the interactions
  with other components (see \defn{im}).  However, in the context
  of this paper we can simplify by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, e} q'$ and $q \goesto{a, g, e}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

@@

  We write $\alpha \prec_\pi \beta$ for $(\alpha, \beta) \in \pi$ and 
  denote $\filter{\alpha} \bydef{=} \setdefb{\beta \in (2^P \times
    \guards{V} \times \exprs{V})}{\alpha \prec_\pi \beta}$ the set of
  interactions that have higher priority than $\alpha$.
\end{definition}

\begin{definition}[Application of priority]
  \label{defn:priority:sem}
  The application of a priority model $\pi$ over $(V, P, \export)$
  to a component $B = (Q, q^0, V, \val{0}{}, P, \export, \goesto{})$
  is the component $\pi(B) = (Q, q^0, V, \val{0}{}, P, \export,
  \goesto[\pi]{})$, where $\goesto[\pi]{}$ is the minimal transition
  relation satisfying the rule
%
  \begin{equation}
    \label{eq:priority:sem}
%
    \infer{    
      q \goesto{a, g, e} q'
      \and
      \textstyle
      g' = g \land \bigwedge_{(b, h, f) \in \filter[q]{\alpha}} \lnot h
    }{
      q \goesto[\pi]{a, g', e} q'
    }\,, 
  \end{equation}
%
  where $\filter[q]{\alpha} \bydef{=} \filter{\alpha} \cap
  \setdefb{\beta \in (2^P \times \guards{V} \times \exprs{V})}{q
    \goesto{\beta}}$.
  \todoSBin{Add the intuitive explanation}
\end{definition}

@@

\begin{note}
  \label{rem:maxprogress}
  Below, we will assume that the maximal progress priority $\mu$ is
  applied after each composition with an interaction model.
\end{note}

@@

\begin{definition}[Application of an architecture]
  \label{defn:arch:application}
  Let $A = (\cC, V_A, P_A, \export[A], \Gamma)$ be an architecture and let $\cB$
  be a set of components, such that
%
  \begin{align}
    \bigcup_{B \in \cB} V_B \cap \bigcup_{C \in \cC} V_C = \emptyset\,,
    &&
    V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B\,,
    \\
    \bigcup_{B \in \cB} P_B \cap \bigcup_{C \in \cC} P_C = \emptyset\,,
    &&
    P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B\,,
  \end{align}
%
  \hl{and $\export[A](p) = V_A \cap \export[B](p)$, for any $B \in \cB$
  and $p \in P_A \cap P_B$.}
%
  The \emph{application of the architecture $A$ to the set of
  components $\cB$} is the component $ A(\cB) \bydef{=}
  \mu\bigl((\IMextend{\Gamma}{P})(\cC \cup \cB)\bigr)$, where
%
  $\IMextend{\Gamma}{P}$ is the \emph{extension} of the interaction
  model $\Gamma$ to the set of ports $P$, \ie the interaction model
  over $(V, P, \export)$, with
  \[
  \export(p) = 
  \begin{cases}
    \export[C](p), & \text{for all } C \in \cC, p \in P_C,
    \\
    \export[B](p), & \text{for all } B \in \cB, p \in P_B,
  \end{cases}
  \]
  defined by
%
  \begin{equation}
    \label{eq:im:extension}
    \IMextend{\Gamma}{P} \bydef{=}
    \setdefb{
      (a \cup a', g, e)
    }{
      (a, g, e) \in \Gamma, a' \subseteq P \setminus P_A
    }
    ;
  \end{equation}
%
  $\mu$ is the maximal progress priority model.
\end{definition}

@@

\begin{definition}[Composition of architectures]
  \label{defn:arch:composition}
  Let $A_i = (\cC_i, V_{A_i}, P_{A_i}, \export[A_i], \Gamma_i)$, for $i = 1,2$,
  be two architectures.  The \emph{composition} of $A_1$ and
  $A_2$ is the architecture $A_1 \arcomp A_2 = (\cC_1 \cup \cC_2,
  V_{A_1} \cup V_{A_2}, P_{A_1} \cup P_{A_2}, \export[A_1] \cup \export[A_2], \Gamma)$, where
%
  \begin{equation}
    \label{eq:arch:composition}
    \Gamma = \setdefb{
      (a, g^1 \land g^2, e^1 \expmix e^2) 
    }{
      (a \cap P_{A_i}, g^i, e^i) \in \Gamma_i,
      \text{ for } i = 1,2
    }
    \,.
  \end{equation}
\end{definition}

\begin{proposition}[Properties of $\arcomp$]
  \label{prop:arcomp:nice}
  Architecture composition $\arcomp$ is commutative and
  associative; it is idempotent if all coordinating components
  are deterministic; $A_{id} = \bigl(\emptyset, \emptyset,
  \emptyset, \emptyset, \{(\emptyset, \true, \noop)\}\bigr)$ is its neutral element, \ie for
  any architecture $A$, we have $A \oplus A_{id} \arequiv A$.
  Furthermore, for any component $B$, we have $A_{id}(B) = \mu(B)$.
\end{proposition}
%
\begin{proof}[Sketch of the proof]
  Commutativity and associativity follow from the corresponding
  properties of set union, Boolean conjunction and
  \todoSB{Check ``semi''}\hl{semilattice meet}.
%
  Suppose we have two architectures $A \arequiv A'$.  This does not
  necessarily mean that their sets of coordinating components
  coincide.  However, if all the involved coordinating components
  are deterministic, then, in any state of $(A \arcomp A')(\cB)$,
  both architectures will impose the same restrictions, enabling
  the same interactions between the coordinating and operand
  components.  Hence, we have $(A \arcomp A')(\cB) = A(\cB) =
  A'(\cB)$.  Since this holds for any set of components $\cB$, we
  conclude that $A \arcomp A' \arequiv A \arequiv A'$.
%
  The properties of $A_{id}$ follow immediately from the
  definitions of architecture application and composition.
\end{proof}

The last statement of \prop{arcomp:nice} highlights a subtle point:
$A_{id}$ is the neutral element of the $\arcomp$ operator\mdash it is
not the identity operator on components.

@@

\begin{definition}[Upwards compatibility]
  \label{defn:up:compat}
  \todoSB{Summarise in English and move to the appendix}
%
  Consider two architectures $A_i = (\cC_i, V_{A_i}, P_{A_i},
  \export[A_i], \Gamma_i)$, for $i = 1,2$.
%
  We say that $A_1$ and $A_2$ are \emph{upwards compatible} iff, for
  any two interactions $(a_i, g_i, e_i) \in \Gamma_i$, such that $a_1
  \cap P_{A_2} = a_2 \cap P_{A_1}$ (\ie they agree on the common ports
  of the two architectures), we have 1)~for any interaction $(b_1,
  h_1, f_1) \in \Gamma_1$, such that $a_1 \varsubsetneq b_1$, there
  exists an interaction $(b_2, h_2, f_2) \in \Gamma_2$, such that $a_2
  \subseteq b_2$, $b_1 \cap P_{A_2} = b_2 \cap P_{A_1}$ and the 
  implication  
%
  \begin{multline*}
    \left(
    \exists (q^1_C)^{C \in \supp{b_1} \cap \cC_1}:
    \exists (q^2_C)^{C \in \supp{a_2} \cap \cC_2}:
    \bigwedge_{
      \begin{multlined}
        \scriptstyle
        C \in \supp{b_1} \cap \cC_1 \cap
        \\
        \scriptstyle
        \supp{a_2} \cap \cC_2
      \end{multlined}
    }
    q^1_C = q^2_C
    \ \land
    \right.
    \\
    \begin{aligned}
      \exists \bigl(q^1_C \goesto{b_1 \cap P_C, g^{b_1}_C, e^{b_1}_C}\bigr)^{C \in \supp{b_1} \cap \cC_1}:
      & \quad \left.
      h_1 \land \bigwedge_{C \in \supp{b_1} \cap \cC_1} g^{b_1}_C
      \ \land \right.
      \\
      \exists \bigl(q^1_C \goesto{a_1 \cap P_C, g^{a_1}_C, e^{b_1}_C}\bigr)^{C \in \supp{a_1} \cap \cC_1}:
      & \quad \left.
      g_1 \land \bigwedge_{C \in \supp{a_1} \cap \cC_1} g^{a_1}_C
      \ \land \right.
      \\
      \exists \bigl(q^2_C \goesto{a_2 \cap P_C, g^{a_2}_C, e^{a_2}_C}\bigr)^{C \in \supp{a_2} \cap \cC_2}:
      & \quad
      \left.
      g_2 \land \bigwedge_{C \in \supp{a_2} \cap \cC_2} g^{a_2}_C
      \right) \implies 
    \end{aligned}
  \end{multline*}
%
  \begin{multline*}
    \left(
    \exists (q^2_C)^{C \in (\supp{b_2}\setminus\supp{a_2}) \cap \cC_2}:
            {\color{white}\bigwedge_{C \in \cC_1}}\right. % Fake bigwedge to make the parenthesis the right size
            \\
            \left.
            \exists \bigl(q^2_C \goesto{b_2 \cap P_C, g^{b_2}_C, e^{b_2}_C}\bigr)^{C \in \supp{b_2} \cap \cC_2}:
            h_2 \land
            \bigwedge_{C \in \supp{b_2} \cap \cC_2} g^{b_2}_C
            \right)
  \end{multline*}
%  
  holds for all valuations $\val{}{1} : V_{A_1} \rightarrow \data$
  and $\val{}{2} : V_{A_2} \rightarrow \data$; and 2)~symmetrically
  for any interaction $(b_2, h_2, f_2) \in \Gamma_2$, such that $a_2
  \varsubsetneq b_2$.
\end{definition}

\begin{theorem}[Preserving enforced properties]
  \label{thm:combining}
  Let $\cB$ be a set of components; let $A_i = (\cC_i, V_{A_i},
  P_{A_i}, \export[A_i], \Gamma_i)$, for $i = 1,2$, be two upwards compatible architectures
  enforcing on $\cB$ the properties $\Phi_1$ and $\Phi_2$
  respectively.  The composition $A_1 \arcomp A_2$ enforces on
  $\cB$ the property $\Phi_1 \land \Phi_2$.
\end{theorem}

@@

\begin{lemma}
  \label{lem:onlyone}
  Let $A = (\cC, V_A, P_A, \export[A], \Gamma)$ be an architecture and denote
  by $\Gamma_\cC \bydef{=}
%
  \setdef{
    (a \cap P_\cC, \true, \noop)
  }{
    (a, g, e) \in \Gamma
  }$, with $P_\cC = \bigcup_{C \in \cC} P_C$,
%  
  the projection of $\Gamma$ onto the coordinating components of
  $A$.  Consider the architecture $A' = (\{C'\}, V_A, P_A, \export[A],
  \Gamma)$, where $C' = \Gamma_\cC(\cC)$.  For any set of
  components $\cB$, satisfying the conditions of
  \defn{arch:application}, we have
  $\semopen{A(\cB)} = \semopen{A'(\cB)}$.
\end{lemma}
%
\begin{proof}
  First of all, notice that, by \defn{im:sem},
  $V_{C'} = \bigcup_{C \in \cC} V_C$ and $P_{C'} = P_\cC = \bigcup_{C \in \cC} P_C$.
  The conditions of \defn{arch} are satisfied and $A'$ is indeed
  an architecture.  Furthermore, $\cB$ satisfies the conditions
  of \defn{arch:application} \wrt $A'$.  Hence, the component
  $A'(\cB)$ is well defined.

  Clearly the state spaces and initial states %% and interfaces
  of both LTSs $\semopen{A(\cB)}$ and $\semopen{A'(\cB)}$
  coincide.  Thus, we only have to prove that so do the
  transition relations.  Let us assume that
  $\cC = C_i^{i \in I}$ and $\cB = B_j^{j \in J}$.
  We will use $q_i, q_i'$ to denote the states of
  $C_i$ and $q_j, q_j'$ to denote the states of $B_j$,
  and similarly for the valuations of variables.

  By \defn{comp:semantics},
%
  \begin{equation}
    \label{eq:lem1:trans:sem}
    (q_i, \val{}{i})^{i \in I} (q_j, \val{}{j})^{j \in J}
    \goesto{a, \val[\tilde]{}{}}
    (q_i', \val{\prime}{i})^{i \in I}
    (q_j', \val{\prime}{j})^{j \in J}
  \end{equation}
%  
  is a transition in $\semopen{A(\cB)}$ (\resp
  $\semopen{A'(\cB)}$) iff
%
  \begin{equation}
    \label{eq:lem1:trans}
    q_i^{i \in I} q_j^{j \in J}
    \goesto{a, g', e'}
    (q_i')^{i \in I} (q_j')^{j \in J}
    \,,
  \end{equation}
%
  is a transition in $A(\cB)$ (\resp $A'(\cB)$) and
%
  \begin{mathpar}
    \val{i \in I}{i} \val{j \in J}{j}
    \models g'
    \,,
    \and
    (\val{\prime}{i})^{i \in I} (\val{\prime}{j})^{j \in J}
    = \val[\tilde]{}{}[e']
    \,,
    \and
    \valdiff{\val{i \in I}{i} \val{j \in J}{j}}{
      \val[\tilde]{}{}} \subseteq \export(a)       
  \end{mathpar}
%
  (with $\export$ defined as in \defn{arch:application}).
